Charity During Crackdown: Analyzing the Impact of State Repression of NGOs on Philanthropy

### 2019-03-12 manuscript/manuscript.Rmd
More shrinking

In January 2011, the Egyptian uprisings presented an opportunity for domestic non-governmental organizations (NGOs) to leverage their knowledge and expertise to transform Egypt into a democratic state. However, just as NGOs began to take advantage of new levels of public engagement and civic freedoms, successive Egyptian governments passed laws restricting funding and advocacy by civil society. Subsequently, international donor agencies cut support for Egyptian NGOs out of concern that they would be accused of offering them illegal support [@Brechenmacher:2017]. Egypt is not alone in its non-violent crackdown on NGOs. India, Bolivia, Hungary, China, Russia, and others have required that funds to NGOs be routed through state-owned banks, mandated that funds cannot be used for political purposes, or prohibited NGOs from accepting any funds from foreign sources [@Carothers:2015; @DPR:2015].

These restrictive laws have had tangible consequences for NGOs worldwide. Official aid flows have decreased substantially in countries that have imposed restrictive anti-NGO laws [@ChaudhryHeiss:2018; @DupuyPrakash:2018], and large foundations and government agencies have both reduced their contributions to NGOs in restrictive countries [@Brechenmacher:2017]. In response to this dramatic shift in the NGO funding landscape, organizations have increasingly turned to individual private donors to offset the loss of governmental and foundation funding [@BanksHulmeEdwards:2015]. In 2016, U.S. private philanthropy for international NGOs amounted to \$22.06 billion [@givingusa:2017].

Donor location poses a challenge for NGOs, however. Donations from local host-country *domestic* donors in repressive countries are unreliable and inconsistent. Citizens in many countries in the Global South often cannot effort to support local NGOs, may lack a culture of philanthropic giving, or may prefer to channel funds to groups working on non-contentious issues such as health and education [@DPR:2015; @RPC:2016; @Brechenmacher:2017]. Even when donors in repressive countries wish to donate to NGOs working in contentious areas, they are often deterred by dismal tax incentives, threats to take away business permits, or have a general fear of retribution [@HudsonInstitute:2015]. In some authoritarian contexts, legal or structural changes make it impossible or counter-productive to donate [@Baoumi:2016]. Accordingly, NGOs working abroad have increasingly sought out funds from *foreign* individual donors in countries with institutions and regulations better suited for philanthropy. Further, given exchange rates and wealth differentials, it is easier for local NGOs to raise modest amounts of money abroad, rather than investing a large amount of resources and manpower to raise similar sums locally [@RonKairePandya:2018]. Given increased crackdown on NGOs across the world, how do individual donors in the U.S. feel about donating to legally besieged NGOs abroad? Do legal restrictions on NGOs influence donors' decision to donate?

@@

We investigate these questions by exploring the structural determinants of private philanthropy. Substantial research has looked at the private motivations of donors, including altruism, reputational benefits, and alignment with personal values [@BilodeauSlivinski:1997; @Wiepking:2010; @BekkersWiepking:2011]. The decision to donate to an NGO is rarely rational, however [@Tremblay-BoirePrakash:2017]. In theory, perfectly rational donors should scrutinize each aspect of an NGO's organizational structure and programmatic performance prior to donation. Instead, research has shown that donors rely on shortcuts, signals, and heuristics to determine the trustworthiness of NGOs. These heuristics are commonly rooted in donor perceptions of organizational characteristics—donors make cursory judgments about an organization's issue area, mission, vision, and values, and seek out supplementary information from friends, family, and acquaintances [@SzperPrakash:2011; @Sloan:2009]. In addition to organizational characteristics that signal an NGO's trustworthiness and deservingness we argue that deeper structural factors—including the nationality of the targets of NGO programs and whether an organization works abroad—can also serve as important heuristics in the decision to engage in philanthropy [@Rajan:2009; @CasaleBaumann:2015; @KnowlesSullivan:2017; @Tremblay-BoirePrakash:2017].

We propose that nature of the legal and regulatory relationship between international NGOs and their host governments can also serve as a structural signal to donors of NGO trustworthiness and deservingness. NGO-related legislation is not inherently restrictive. Governments frequently use regulations to routinize the behavior of NGOs—regulations produce convergent practices [@BloodgoodTB, 145], prevent malfeasance that threatens to undermine confidence in the entire NGO community [@GibelmanGelman; @ONeill2009], and help improve NGO access to the policy-making process [@Teets:2014]. Here, however, we focus on the impact of restrictive regulations or anti-NGO laws. Unlike regulations aimed at standardizing state–NGO relations, these laws are designed to limit the NGO community by creating deliberate barriers to their entry, funding, and advocacy. These laws can even limit organizations who comply with regulations regarding registration, auditing, and transparency.

@@

> **H~1a~**:  If NGOs face legal crackdowns abroad, individual donors will be more likely to donate to them.
> 
> **H~1b~**:  If NGOs face legal crackdowns abroad, individual donors will donate more to them.

@@

In contrast, NGOs working on more contentious issues like human rights, election monitoring, anti-corruption, and youth movements can pose political challenges to regimes and are less compatible with their host governments. Accordingly, these NGOs are often viewed as politically motivated outsiders, due to their advocacy against state-sanctioned abuses and efforts to changing discriminatory policies and attitudes that threaten the rights of marginalized and unpopular groups [@Crowther:2015]. Thus, they run a higher risk of getting shut down [@DPR:2015], increasing the possibility of donors' resources being wasted. Their programmatic effectiveness is also harder to measure and quantify. Further, donors may believe that it is more difficult to propagate human rights in the target country's society, which limits NGOs' ability to achieve their goals.

@@

> **H~2a~**:  If legally besieged NGOs work on humanitarian issues, individual donors will be more likely to donate to them.
> 
> **H~2b~**:  If legally besieged NGOs work on humanitarian issues, individual donors will donate more to them.

The *source of NGO funding* can also influence donor behavior. Government grants to NGOs can crowd out private funding and significant reductions in fundraising from private donors [@AndreoniPayne:2003]. In addition, donors may feel less inclined to contribute since the NGO does not need (or might not even typically accept) their donation, and since donors might perceive that their tax money already funds the organization. Further, government funding for NGOs can also potentially signal a lack of independence from the government. Organizations that receive substantial government funding frequently avoid programming that would question the government even if it is in line with their mission [@Stroup:2012]. In contrast, donors who see that organizations are privately funded will feel that they can also contribute and help.

Funding sources also have an effect on anti-legal crackdowns on NGOs (see @fig:causal-path), given that funding restrictions are one of the most common forms of anti-NGO regulations [@DPR:2015]. As such, the combination of legal crackdown and funding may reshape donor perceptions of deservingness. Donors could punish government-funded NGOs that face restrictions, since legal trouble could further fuel donor mistrust, and since anti-NGO crackdowns are commonly based on fears of foreign intervention in domestic politics. Donors thus are more likely to donate more to NGOs that rely on private funding:

> **H~3a~**:  If legally besieged NGOs do not receive substantial funding from government sources, individual donors will be more likely to donate to them.
>
> **H~3b~**: If legally besieged NGOs do not receive substantial funding from government sources, individual donors will donate more to them.

@@

These results give partial support to H~3a~. The source of funding has little influence on the propensity to donate to humanitarian NGOs, but on average, donors are substantially more likely to give to privately-funded human rights NGOs than government-funded organizations. However, the effect of crackdowns reverses across conditions, as donors appear to both punish government-funded human rights NGOs that face legal difficulties and rally behind privately-funded NGOs facing those same hurdles. Many of the respondents described their reasoning for removing support for besieged government-funded organizations, explaining that "[i]f the non-profit works in countries where non-profit activities are heavily restricted there's a good amount of waste and shortfalls in their activities and ability to assist others,"[^3154467] and that accordingly, they would be less likely to contribute. Others explained that they would decrease their donations to government-backed organizations because of the bad optics of legal crackdown, with one stating outright that "I'm not exactly sure what they did was right or wrong, I think it just seems bad, so I'm less likely to want to donate to them."[^8425616]

While donors on average appear to punish government-funded human rights organizations, they increase their support to privately funded NGOs. Many respondents explained their motivation for being more likely to donate in moralistic, normative terms, reasoning that "[s]ince I was small child my parents thought [sic] me to help others,"[^6009603] or "I would want to do the right thing."[^9526638] Another explained "[t]hey [IRC] are doing good work in countries where it is tough for groups like them to operate and they need all the help they can get."[^3025921] Individual donors thus seem to be more willing to support besieged human rights organizations when they are unencumbered by government funds.

@@

This trend also holds true when accounting for the issue area of the NGO. As seen in @fig:amount-diffs(B–C), the crackdown condition elicits higher donations for both the human rights and humanitarian assistance conditions, though with varying levels of significance. Emphasizing legal crackdown increased donations to human rights NGOs by `r md_dollar(filter(amount, title == "Human rights")$mu_diff)` above the no crackdown condition, but with a lower probability of significance (`r filter(amount, title == "Human rights")$p_neq0`). In contrast, crackdowns significantly boosted donations to humanitarian assistance NGOs by `r md_dollar(filter(amount, title == "Humanitarian assistance")$mu_diff)`, a `r percent_0(filter(amount, title == "Humanitarian assistance")$pct_change)` increase beyond the control condition (`r filter(amount, title == "Humanitarian assistance")$p_neq0`). We thus find support for H~2b~: donors give more to besieged humanitarian NGOs. In general, informing donors about crackdowns causes them to give more to humanitarian NGOs facing legal difficulties abroad. We also find evidence for H~3b~. Donors give `r md_dollar(filter(amount, title == "Private")$mu_diff)` more (a `r percent_0(filter(amount, title == "Private")$pct_change)` increase) to legally besieged NGOs that are privately funded (`r filter(amount, title == "Private")$p_neq0`), while the crackdown condition has little effect on the amount donated to government-funded NGOs (`r filter(amount, title == "Government")$p_neq0`).

Finally, combining issue and funding sources adds more texture to these results. We previously found that donors seem to (1) punish government-funded human rights NGOs and (2) reward privately-funded human rights NGOs when they run into legal issues. This finding also applies somewhat to the amount donors give. Respondents donated `r md_dollar(filter(amount, title == "Human rights | Private")$mu_diff)` more to besieged privately-funded human rights NGOs compared to the no crackdown control condition (`r filter(amount, title == "Human rights | Private")$p_neq0`), increasing their donations by `r percent_0(filter(amount, title == "Human rights | Private")$pct_change)`. Anecdotally, many of the respondents explained that they chose to give additional money to privately-funded human rights NGOs precisely because "the country limits non-profits."[^13685] The punishment mechanism found previously does not translate to the amount donated—while there is a decrease of `r md_dollar(abs(filter(amount, title == "Human rights | Government")$mu_diff))` in average donations to government-funded human rights NGOs facing crackdown, the chance this difference is less than zero is lower than our threshold (`r filter(amount, title == "Human rights | Government")$p_neq0`).

Emphasizing legal difficulties increased donations to government-funded humanitarian assistance NGOs by `r percent_0(filter(amount, title == "Humanitarian assistance | Government")$pct_change)`, or `r md_dollar(abs(filter(amount, title == "Humanitarian assistance | Government")$mu_diff))` (`r filter(amount, title == "Humanitarian assistance | Government")$p_neq0`). The crackdown condition caused this increase despite some notable hesitation from respondents, many of whom felt that government-backed NGOs do not need or deserve private donations. Respondents reasoned since IRC "is already receiving funding from governments, so funding from private individuals doesn't seem as necessary. I'd prefer to give my money to an organization that is primarily run by private donations"[^6486084] or because "[m]y tax paying dollars go to the government and if the government is helping to fund it then they are in turn getting some of my money."[^9509936] This hesitation was not universal, however, and in aggregate most respondents chose to give substantially more. Many echoed this respondent, who identified the crackdown condition as one of the reasons for giving the full hypothetical \$100: "I believe the work of such organizations has an urgency that affects many refugees and their families. If they can function without being affected by the harsh laws against non-profits, I would want them to have as much help as possible to do their humanitarian work."[^1097165]

In contrast, crackdowns had no substantial effect on the amount donated to privately-funded humanitarian NGOs (`r filter(amount, title == "Humanitarian assistance | Private")$p_neq0`). None of the respondents in these conditions mentioned funding sources when justifying their giving, and many explained that they would not give solely on the basis of the legal crackdown, including one who stated that they "would keep all \$100 because it would not even go to helping humanitarian efforts due to the new laws."[^8995923] It is possible that donors see government-funded humanitarian NGOs as more legitimate and more capable of handling difficult legal restrictions than privately-funded organizations. For instance, donors might assume that an international NGO that receives substantial money from USAID would ostensibly have the backing and support of the U.S. government when facing legal challenges, while an NGO funded by small donors would be less able to confront such challenges. Further research is needed to probe this trend.

@@

Over the last three decades, dozens of countries have used legal restrictions to limit international NGO funding and programming. Between the lack of a local philanthropic culture, an increasingly authoritarian approach that restricts space for domestic philanthropy, and the withdrawal of official governmental aid, many international NGOs have turned to foreign private donors for their funding. But how do these donors respond to anti-NGO legal restrictions, and how do an organization's issue area and funding sources interact with these restrictions? In this study, propose that legal crackdowns on NGOs can serve as a heuristic or signal to donors of the deservingness and efficiency of NGO programming. We find that crackdowns, funding, and issue area do not directly influence the *likelihood* of donation on their own. Donors are more likely to give to privately funded NGOs facing legal crackdown, however, and are more likely to do so when those organizations focus on human rights issues. Donors appear to punish government-funded human rights organizations that face crackdown and are less likely to support them. Crackdowns do, however, influence the *amount* that donors give. Respondents indicated a willingness to donate substantially more to NGOs facing legal restrictions, and this effect persists when mediated by issue area and funding—respondents give more to restricted humanitarian NGOs and privately funded NGOs. For human rights NGOs, those funded through private donations see increased donations under legal crackdown, while donors again appear to punish those that rely on government funding.

This study makes two important contributions. While recent work has looked at how state repression of NGOs affects official aid flows [@ChaudhryHeiss:2018; @DupuyPrakash:2018], there is little research on how such repression affects private aid flows. Further, prior literature on the determinants of donation to NGOs predominantly focuses on individual-level factors. Our study measures important structural determinants that influence giving and shows how these factors interact with organizational attributes to signal NGO deservingness. Additional research should determine the extent to which these preferences translate into action—here, we have only asked respondents to make a hypothetical donation—as well as test the generalizability of our findings to other issue areas.

Our results have several implications for policy and future research. NGOs may actually benefit from publicizing when they are targets of government crackdown. While this may not be enough to sway the average U.S. citizen, our results show that communicating this information can convince already-likely donors to donate more to besieged groups. Our study also highlights the importance of framing issue areas. {{TODO: Donors are likely to view legally besieged human rights NGOs with suspicion. This framing is critical, as previous research on human rights micromobilization has shown that the way people process information can affect their increased willingness to help others in need [@SmallLoewensteinSlovic:2007; @Slovic:2007]. However, human rights groups undertake a variety of programs and missions, and communicating the nature of their work and specifying the main target of the organization's efforts can allay the public's fears.}} Finally, if NGOs signal to donors that receive a majority of their funding from non-governmental sources, they may be able to convince average donors that their contributions might help the NGO resist the crackdown. Moreover, since private funders tend to be less interested in the domestic politics of recipient countries compared to governmental funders, relying on private donations is less likely to attract the ire of authoritarian governments fearful of foreign intervention.

With increasing government repression of NGOs, these groups can no longer rely on a single funding source. However, these new and challenging scenarios create an opportunity for NGOs to rethink how to address different audiences and create a robust market for philanthropy towards besieged NGOs abroad.

>>>>

In January 2011, the Egyptian uprisings presented an opportunity for domestic non-governmental organizations (NGOs) and civil society actors to transform Egypt into a democratic state. However, just as NGOs began to take advantage of new levels of civic freedoms, successive Egyptian governments passed laws restricting these groups from receiving foreign funding or engaging in political advocacy. Egypt is not alone in its non-violent crackdown on NGOs. India, Bolivia, Hungary, China, Russia, and others have mandated that funds cannot be used for political purposes, or prohibited NGOs from accepting any funds from foreign sources [@Carothers:2015; @DPR:2015].

These restrictive laws have tangible consequences for NGOs worldwide. Official aid flows have decreased substantially in countries that have imposed restrictive anti-NGO laws [@ChaudhryHeiss:2018; @DupuyPrakash:2018], and large foundations and government agencies have both reduced their contributions to NGOs in restrictive countries [@Brechenmacher:2017]. In response to this dramatic shift in the NGO funding landscape, organizations have increasingly turned to individual private donors to offset the loss of governmental and foundation funding [@BanksHulmeEdwards:2015]. In 2016, U.S. private philanthropy for international NGOs amounted to \$22.06 billion [@givingusa:2017].

Donor location poses a challenge for NGOs, however. Donations from *domestic* donors in repressive countries are unreliable and inconsistent. Citizens in many countries in the Global South often cannot support local NGOs, may lack a culture of philanthropic giving, or may channel funds to groups working on non-contentious issues such as health and education [@DPR:2015; @RPC:2016; @Brechenmacher:2017]. Even when donors in repressive countries prefer to donate to NGOs working in contentious areas, they may be deterred by dismal tax incentives, threats to take away business permits, or have a general fear of retribution [@HudsonInstitute:2015]. In some authoritarian contexts, legal or structural changes make it impossible or counter-productive to donate [@Baoumi:2016]. Accordingly, NGOs have increasingly sought out funds from *foreign* individual donors in countries with institutions and regulations better suited for philanthropy. Further, given exchange rates and wealth differentials, it is easier for local NGOs to raise modest amounts of money abroad, rather than investing a large amount of resources and manpower to raise similar sums locally [@RonKairePandya:2018]. Given increased crackdown on NGOs across the world, how do individual donors in the U.S. feel about donating to legally besieged NGOs abroad? Do legal restrictions on NGOs influence donors' decision to donate?

@@

We investigate these questions by exploring the structural determinants of private philanthropy. Substantial research has looked at the private motivations of donors, including altruism, reputational benefits, and alignment with personal values [@BilodeauSlivinski:1997; @BekkersWiepking:2011]. The decision to donate to an NGO is rarely rational, however [@Tremblay-BoirePrakash:2017]. In theory, perfectly rational donors should scrutinize each aspect of an NGO's organizational structure and programmatic performance prior to donation. Instead, research has shown that donors rely on shortcuts, signals, and heuristics to determine the trustworthiness of NGOs. These heuristics are commonly rooted in donor perceptions of organizational characteristics—donors make cursory judgments about an organization's issue area, mission, vision, and values, and seek out supplementary information from friends, family, and acquaintances [@SzperPrakash:2011; @Sloan:2009]. In addition to organizational characteristics that signal an NGO's trustworthiness and deservingness we argue that structural factors—including the nationality of the targets of NGO programs and whether an organization works abroad—can also serve as important heuristics in the decision to engage in philanthropy [@CasaleBaumann:2015; @KnowlesSullivan:2017; @Tremblay-BoirePrakash:2017].

We propose that nature of the legal and regulatory relationship between international NGOs and their host governments can serve as a signal of NGO trustworthiness and deservingness. NGO-related legislation is not inherently restrictive. Governments frequently use regulations to routinize the behavior of NGOs—regulations produce convergent practices [@BloodgoodTB, 145] and help improve NGO access to the policy-making process [@Teets:2014]. Here, however, we focus on the impact of restrictive regulations or anti-NGO laws. Unlike regulations aimed at standardizing state–NGO relations, these laws are designed to limit the NGO community by creating deliberate barriers to their entry, funding, and advocacy. These laws can even limit organizations who comply with regulations regarding registration, auditing, and transparency.

@@

> **H~1a~**:  If NGOs face legal crackdowns abroad, then individual donors will be more likely to donate to them.
> 
> **H~1b~**:  If NGOs face legal crackdowns abroad, then individual donors will donate more to them.

@@

In contrast, NGOs working on more contentious issues like human rights, election monitoring, anti-corruption, and youth movements can pose political challenges to regimes. Therefore, these INGOs are often viewed as politically motivated outsiders and they run a higher risk of getting shut down [@DPR:2015], increasing the possibility of donors' resources being wasted. Due to the nature of the political and social issues they work on, their programmatic effectiveness is also harder to measure and quantify [@Bush:2015]. Thus, donors may believe that it is more difficult to improve respect for human rights in the target country, which limits NGOs' ability to achieve their goals.

@@

> **H~2a~**:  If legally besieged NGOs work on humanitarian issues, then individual donors will be more likely to donate to them.
> 
> **H~2b~**:  If legally besieged NGOs work on humanitarian issues, then individual donors will donate more to them.

The *source of NGO funding* can also influence donor behavior. Government grants to NGOs can often crowd out private funding [@AndreoniPayne:2003]. In addition, donors may feel less inclined to contribute since the NGO does not need (or might not even typically accept) their donation, and since donors might perceive that their tax money already funds the organization. Further, government funding for NGOs can also potentially signal a lack of independence from the government. Organizations that receive substantial government funding frequently avoid programming that would question the government even if it is in line with their mission [@Stroup:2012]. In contrast, donors who see that organizations are privately funded will feel that they can also contribute and help.

Funding sources also have an effect on anti-legal crackdowns on NGOs (see @fig:causal-path), given that funding restrictions are one of the most common forms of anti-NGO regulations [@DPR:2015]. As such, the combination of legal crackdown and funding may reshape donor perceptions of deservingness. Donors could punish government-funded NGOs that face restrictions, since legal trouble could further fuel private donor mistrust. Donors thus are more likely to donate more to NGOs that rely on private funding:

> **H~3a~**:  If legally besieged NGOs do not receive substantial funding from government sources, then individual donors will be more likely to donate to them.
>
> **H~3b~**: If legally besieged NGOs do not receive substantial funding from government sources, then individual donors will donate more to them.

@@

These results give partial support to H~3a~. The source of funding has little influence on the propensity to donate to humanitarian NGOs, but on average, donors are substantially more likely to give to privately-funded human rights NGOs than government-funded organizations. However, the effect of crackdowns reverses across conditions, as donors appear to both punish government-funded human rights NGOs that face legal difficulties and rally behind privately-funded NGOs facing those same hurdles. Respondents described their reasoning for the former as "[i]f the non-profit works in countries where non-profit activities are heavily restricted there's a good amount of waste and shortfalls in their activities and ability to assist others."[^3154467] Respondents also explained their motivation for increased support to privately funded NGOs as "[t]hey [IRC] are doing good work in countries where it is tough for groups like them to operate and they need all the help they can get."[^3025921] Individual donors thus seem to be more willing to support besieged human rights organizations when they are unencumbered by government funds.

@@

This trend also holds true when accounting for NGO issue area. As seen in @fig:amount-diffs(B–C), the crackdown condition elicits higher donations for both the human rights and humanitarian NGOs, though with varying levels of significance. Emphasizing legal crackdown increased donations to human rights NGOs by `r md_dollar(filter(amount, title == "Human rights")$mu_diff)` above the no crackdown condition, but with a lower probability of significance (`r filter(amount, title == "Human rights")$p_neq0`). In contrast, crackdowns significantly boosted donations to humanitarian NGOs by `r md_dollar(filter(amount, title == "Humanitarian assistance")$mu_diff)`, a `r percent_0(filter(amount, title == "Humanitarian assistance")$pct_change)` increase beyond the control condition (`r filter(amount, title == "Humanitarian assistance")$p_neq0`). We thus find support for H~2b~: donors give more to besieged humanitarian NGOs. We also find evidence for H~3b~. Donors give `r md_dollar(filter(amount, title == "Private")$mu_diff)` more (a `r percent_0(filter(amount, title == "Private")$pct_change)` increase) to legally besieged NGOs that are privately funded (`r filter(amount, title == "Private")$p_neq0`), while the crackdown condition has little effect on the amount donated to government-funded NGOs (`r filter(amount, title == "Government")$p_neq0`).

Finally, combining issue areas and funding sources adds more texture to these results. We previously found that donors seem to (1) punish government-funded human rights NGOs and (2) reward privately-funded human rights NGOs when they run into legal issues. This finding also applies to the amount donors give. Respondents donated `r md_dollar(filter(amount, title == "Human rights | Private")$mu_diff)` more to besieged privately-funded human rights NGOs compared to the no crackdown control condition (`r filter(amount, title == "Human rights | Private")$p_neq0`), increasing their donations by `r percent_0(filter(amount, title == "Human rights | Private")$pct_change)`. Respondents explained that they chose to give additional money to privately-funded human rights NGOs precisely because "the country limits non-profits."[^13685] The punishment mechanism found previously does not translate to the amount donated—while there is a decrease of `r md_dollar(abs(filter(amount, title == "Human rights | Government")$mu_diff))` in average donations to government-funded human rights NGOs facing crackdown, the chance this difference is less than zero is lower than our threshold (`r filter(amount, title == "Human rights | Government")$p_neq0`).

Emphasizing legal difficulties increased donations to government-funded humanitarian NGOs by `r percent_0(filter(amount, title == "Humanitarian assistance | Government")$pct_change)`, or `r md_dollar(abs(filter(amount, title == "Humanitarian assistance | Government")$mu_diff))` (`r filter(amount, title == "Humanitarian assistance | Government")$p_neq0`). The crackdown condition caused this increase despite some notable hesitation from respondents, many of whom felt that government-backed NGOs do not need or deserve private donations. Respondents reasoned since IRC "is already receiving funding from governments, so funding from private individuals doesn't seem as necessary. I'd prefer to give my money to an organization that is primarily run by private donations"[^6486084] or because "[m]y tax paying dollars go to the government and if the government is helping to fund it then they are in turn getting some of my money."[^9509936] This hesitation was not universal, however, and in aggregate most respondents chose to give substantially more. Many echoed this respondent, who identified the crackdown condition as one of the reasons for giving the full hypothetical \$100: "I believe the work of such organizations has an urgency that affects many refugees and their families. If they can function without being affected by the harsh laws against non-profits, I would want them to have as much help as possible to do their humanitarian work."[^1097165]

In contrast, crackdowns had no substantial effect on the amount donated to privately-funded humanitarian NGOs (`r filter(amount, title == "Humanitarian assistance | Private")$p_neq0`). No respondent in these conditions mentioned funding sources in their justification, and many explained that they would not give solely because of the crackdown and that they "would keep all \$100 because it would not even go to helping humanitarian efforts due to the new laws."[^8995923] It is possible that donors see government-funded humanitarian NGOs as more legitimate and more capable of handling difficult legal restrictions than privately-funded organizations. For instance, donors might assume that an international NGO that receives substantial money from USAID would ostensibly have the support of the U.S. government when facing legal challenges, while an NGO funded by small donors would be less able to confront such challenges. Further research is needed to probe this trend.

@@

Over the last three decades, several countries have used legal restrictions to limit international NGO funding and programming. Between the lack of a local philanthropic culture, an increasingly authoritarian approach that restricts space for domestic philanthropy, and the withdrawal of official governmental aid, many international NGOs have turned to foreign private donors for their funding. How do these donors respond to anti-NGO legal restrictions, and how do an organization's issue area and funding sources interact with these restrictions? In this study, we propose that legal crackdowns on NGOs can serve as a heuristic or signal to donors of the deservingness and efficiency of NGO programming. We find that crackdowns, funding, and issue area do not directly influence the *likelihood* of donation on their own. Donors are more likely to give to privately funded NGOs facing legal crackdown, however, and are more likely to do so when those organizations focus on human rights issues. Crackdowns do, however, influence the *amount* that donors give. Respondents indicated a willingness to donate substantially more to NGOs facing legal restrictions, and this effect persists when mediated by issue area and funding—respondents give more to restricted humanitarian NGOs and privately funded NGOs. For human rights NGOs, those funded through private donations see increased donations under legal crackdown, while donors again appear to punish those that rely on government funding.

This study makes two important contributions. While recent work has looked at how state repression of NGOs affects official aid flows [@ChaudhryHeiss:2018; @DupuyPrakash:2018], there is little research on how such repression affects private aid flows. Further, prior literature on the determinants of donation to NGOs predominantly focuses on individual-level factors. Our study measures important structural determinants that influence giving and shows how these factors interact with organizational attributes to signal NGO deservingness. Additional research should determine the extent to which these preferences translate into behavior—here, we have only asked respondents to make a hypothetical donation—as well as test the generalizability of our findings to other issue areas.

Our results have several implications for policy and future research. NGOs may benefit from publicizing when they are targets of government crackdown. While this may not be enough to sway the average U.S. citizen, our results show that communicating this information can convince already-likely donors to donate more to besieged groups. Our study also highlights the importance of framing NGO operations. Previous research on human rights micromobilization has shown that the way people process information can affect their increased willingness to help others in need [@Slovic:2007]. Our results show that donors are likely to view legally besieged human rights NGOs with suspicion. However, human rights groups undertake a variety of programs, and communicating the nature of their work and specifying the main target of the organization's efforts can allay the public's fears. Finally, if NGOs signal to donors that receive a majority of their funding from non-governmental sources, they may be able to convince average donors that their contributions might help the NGO resist the crackdown. Moreover, since private funders typically tend to be less interested in the domestic politics of recipient countries compared to governmental funders, relying on private donations is less likely to attract the ire of authoritarian governments fearful of foreign intervention.
